### 2017-11-26 paper/sections/4_dynamic-policies.tex
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\subsection{Trajectory Length}
%
%\subsection{Cloud Techniques: Elastic Search}
%
%What if we re-provision resources in response to events outside the
%application's control, such as a slow Lustre.
%
%\subsection{Future work}
%
%How Much: cache policy from past, regime detection, Belady's Min

@@

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on a single in-memory database node using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}. These strategies are
implemented as the ``when" and ``how much" policies from the Mantle API; the
``where" callback does not make sense for a single node (segments are either in
the cache or they are not). The evaluation metric is the accuracy and runtime
of each strategy; the strategy should be accurate enough so as to sacrifice
negligible performance and fast enough to run as often as we want to detect key
access patterns. The goal of the following sections is not to find an optimal
solution, as this can be done with parameter sweeps for thresholds; rather, we
try to find techniques that work for a range of inputs and system setups.

First we sized the cache according to our system specific knowledge. By
``system", we mean the hardware and software of the storage hierarchy. We look
at request rate, unique keys in a sliding window, and bandwidth capabilities. For
example, we know that LevelDB cannot handle high IO request rates.

% technical details
In the original ParSplice implementation, each cache node uses an unlimited
amount of memory to store segment coordinates. We limit the size of the cache
using an LRU eviction policy, where the penalty for a cache miss is retrieving
the data from the persistent database.  We evict keys (if necessary) at every
operation instead of when segments complete because the cache fills up too
quickly otherwise.

The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The other graph shares the \(y\) axis and shows the trade-off of using a basic
LRU-style cache for different cache sizes, implemented using a ``when" policy
of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
\texttt{servers[whoami]['cachesize']-n}. The error bars are the standard
deviation of 3 runs. 
%\begin{figure}[h] \footnotesize \begin{minted}{lua}
%server[whoami]['cachesize'] > n \end{minted} \end{figure}

@@

% why is the performance lower for smaller caches?
Although the keyspace grows to 150K, a 100K key cache achieves 99\% of the
performance. Decreasing the cache degrades performance and predictability.
Despite the memory savings, our results suggest that a multi-policy cache management strategy
could save even more memory.  The middle plot in
Figure~\ref{fig:methodology-tradeoff} show that a 100K key cache is sufficient
as a static policy but the top graph in Figure~\ref{fig:motivation-regimes}
indicates that the cache size could be much smaller. That graph shows that the
beginning of the run is characterized by many reads to a small set of keys and
the end sees much lower reads per second to a larger keyspace. Specifically, it
shows only about 100 keys as active in the latter half of the run.

After analyzing traces, we see that the 100 key cache is insufficient because
the persistent database cannot service the read-write traffic.  According to
Figure~\ref{fig:motivation-regimes}, the read requests arrive at 750 reads per
second in addition to the writes that land in each tier (about 300 puts/second,
some redundant). This traffic triggers a LevelDB compaction and reads block,
resulting in very slow progress.  Traces verify this hypothesis and show reads
getting backed up as the read/write ratio increases. To recap, small caches
incur too much load on the persistent database  at the beginning of the run but
should suffice after the initial read flash crowd passes because
the keyspace is far less active. This suggests a multi-policy cache management
strategy.

%% Why is this a good idea
%Although ParSplice does not use a distributed file system, its workload is very
%similar because the minima key-value store responds to small and frequent
%requests, which results in hot spots and flash crowds.  Modern distributed file
%systems have found efficient ways to measure, migrate, and partition metadata
%load and have shown large performance gains and
%better scalability~\cite{zheng:pdsw2014-batchfs, zheng:pdsw2015-deltafs,
%grider:pdsw2015-marfs, ren:sc2014-indexfs, patil:fast2011-giga+,
%brandt:msst2003-lh}.  Previous work quantified the speedups achieved with
%Mantle and formalized balancers that were good for file systems.

The plot on the right side of Figure~\ref{fig:methodology-tradeoff} shows the results of using Mantle
to program a multi-policy cache management strategy into ParSplice that switches between
two policies:

@@

The actual policy is shown and described in more detail in the next section in
Figure~\ref{src:lru-dyn}.  We trigger the policy switch at 100K keys to absorb
the flash crowd at the beginning of the run. Once triggered, keys are evicted
to bring the size of the cache down to the threshold.  In the bar chart, the
cache sizes for the \(n\) key limit policy are along the \(x\) axis.

% results: same level of performance can be achieved 
The multi-part policies show better performance than the single \(n\) key
policies. The performance and memory utilization for a 100K key cache size is
the same as the 100K bar in the ``Multi-Part Policy" graph in  Figure~\ref{fig:methodology-tradeoff} but
the rest reduce the size of the keyspace after the read flash crowd.  We see
the worst performance when the engine switches to the 10 key limit policy,
which achieves 94\% of the performance while only using 40KB of memory.

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs vs.  wall-clock time curves down over time; as the nanoparticle grows
it takes longer to generate segments so by the time we reach 2 hours, over 90\%
of the trajectory is already generated.  For (2), the memory footprint is rises
until we reach 100K key switch threshold at 0.4GB, as shown by the
``\(\Delta_1\), Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
can only conclude that this policy only works well for this 2.5 hour run.

% but the result is still valid
Despite these caveats, the result is still valid: we found multi-policy cache 
management strategy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different job permutations.

@@

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.
Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
of keys occurs because molecules are stuck in deep trajectories. Recall that
the in-memory database stores the molecules' EOM minima, which is the smallest
effective energy that a molecule observes during its trajectory. So molecules
stuck in deep trajectories explore the same minima until they can escape to a
new set of states. This exploration of the same set of states is called a
superbasin.

Detecting these superbasins can lead to more effective cache management
strategies because the height of the groups of key accesses is ``how much" of cache
to evict and the width of the groups of key accesses is ``when" to evict values from
the cache.  The zoomed portion of Figure~\ref{fig:keyspace-zoomed} shows how a
single superbasin affects the key accesses. Moving along the \(x\) axis shows
that the number of unique keys accessed over time grows while moving along the
\(y\) axis shows that early keys are accessed more often.  Overall, superbasins
are never re-visited because the simulation only adds molecules; we can never
reach a state with less molecules. This is why keys are never re-accessed.
Despite these patterns, the following characteristics of superbasins make it
hard to detect them:

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that they all
proliferated more parameters that need to be tuned. Furthermore, we find that
many of the metrics do not signal a new set of key accesses, that our metrics
are not accurate enough, and that even if we tune all parameters they still
will not work for all initial conditions.  Below, we indicate which parameters
we need to add for each technique and the value we find to work best, via
tuning and parameter sweeps, for one set of initial conditions.

@@

  \item Statistics: decay on each key counts down until 0 0-valued keys are
evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed large negative slope as signal for new set of key accesses.
``Zero-crossing" for distance between small/large spikes to avoid false
positives, set to 40 seconds; `window size", set to 200 seconds, for the size
of the moving average

@@

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect groups of key
access patterns, which we call ``fans". Iterating backwards through the key
access trace, at each time step we find the lowest key ID and compare against
the lowest key ID we have seen so far (Line 7). We also maintain the top and
bottom of each group of key accesses (Line 13) so we can tell the ``how much"
policy the number of keys to evict (Line 19).  The algorithm iterates backwards
over the key access trace because a change in the minimum value signals a new
group of key accesses. No signal exists iterating left to right, as the maximum
value always increases and the minimum values at the bottom of each group of
key accesses are sparse.  For example, the maximum distance between values
along the bottom edge of the zoomed group of key accesses in
Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
between minimum values for the group of key accesses before is 0 seconds. As a
result of this sparseness, iterating left to right requires a ``window size"
parameter to determine when we think a minimum value will not show up again.
We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
number events, for simplicity because this approach avoids adding new
thresholds for key access pattern detection ({\it e.g.}, space between key
accesses, space between key IDs, and window size of consecutive key accesses).

>>>>

%\subsection{Trajectory Length}
%
%\subsection{Cloud Techniques: Elastic Search}
%
%What if we re-provision resources in response to events outside the
%application's control, such as a slow Lustre.
%
%\subsection{Future work}
%
%How Much: cache policy from past, regime detection, Belady's Min

@@

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on the worker using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}.  The evaluation metric is the
accuracy and runtime of each strategy; the policy should be accurate enough so
as to sacrifice negligible performance and fast enough to run as often as we
want to detect key access patterns.  First we sized the cache according to our
system specific knowledge. By ``system", we mean the hardware and software of
the storage hierarchy. We look at request rate, unique keys in a sliding
window, and bandwidth capabilities. For example, we know that LevelDB cannot
handle high IO request rates.

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph shares the \(y\) axis and shows the trade-off of using a basic
LRU-style cache for different cache sizes, where the penalty for a cache miss
is retrieving the data from the persistent database.  We evict keys (if
necessary) at every operation instead of when segments complete because the
cache fills up too quickly otherwise. This is implemented using a ``when"
policy of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
\texttt{servers[whoami]['cachesize']-n}.  The error bars are the standard
deviation of 3 runs.  Although the keyspace grows to 150K, a 100K key cache
achieves 99\% of the performance. Decreasing the cache degrades performance and
predictability.

%\begin{figure}[h] \footnotesize \begin{minted}{lua}
%server[whoami]['cachesize'] > n \end{minted} \end{figure}

@@

% why is the performance lower for smaller caches?
But the top graph in Figure~\ref{fig:motivation-regimes} suggests that a
smaller cache size should suffice, as only 100 keys seem to be active at any
one time.  It turns out that the unique keys plotted in
Figure~\ref{fig:motivation-regimes} are per second and are not representative
of the actual active keyspace; the number of active keys is larger than 100, as
some keys may be accessed at time \(t_0\), not in \(t_1\), and then again in
\(t_2\). Because the cache is too small, reads and writes hit the persistent
database, where the excessive traffic triggers a LevelDB compaction and reads
block.  To avoid overloading the storage hierarchy with requests, we design a
multi-part policy that switches between:

@@

% results: same level of performance can be achieved 
The key observation is that small caches incur too much load on the persistent
database at the beginning of the run but should suffice after the initial read
flash crowd passes because the keyspace is far less active.  We program Mantle
to trigger the policy switch at 100K keys to absorb the flash crowd at the
beginning of the run. Once triggered, keys are evicted to bring the size of the
cache down to the threshold.  The actual policy is shown and described in more
detail in the next section in Figure~\ref{src:lru-dyn}.  The plot on the right
side of Figure~\ref{fig:methodology-tradeoff} shows the performance/utilization
trade-off of the multi-part policy, where the cache sizes for the \(n\) key
limit policy are along the \(x\) axis.  They show better performance than the
single \(n\) key policies. The performance and memory utilization for a 100K
key cache size is the same as the 100K bar in the ``Multi-Part Policy" graph in
Figure~\ref{fig:methodology-tradeoff} but the rest reduce the size of the
keyspace after the read flash crowd.  We see the worst performance when the
engine switches to the 10 key limit policy, which achieves 94\% of the
performance while only using 40KB of memory.

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows
it takes longer to generate segments so by the time we reach 2 hours, over 90\%
of the trajectory is already generated.  For (2), the memory footprint rises
until we reach 100K key switch threshold at 0.4GB, as shown by the
``\(\Delta_1\), Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
can only conclude that this policy only works well for this 2.5 hour run.

% but the result is still valid
Despite these caveats, the result is still valid: we found a multi-policy cache 
management strategy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different job permutations.

@@

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.  The goal of the following
sections is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
tasks over time (\(x\) axis). The groups of accesses to a subset of keys occurs
because molecules are stuck in deep trajectories. Recall that the cache stores
the molecules' EOM minima, which is the smallest effective energy that a
molecule observes during its trajectory. So molecules stuck in deep
trajectories explore the same minima until they can escape to a new set of
states. This exploration of the same set of states is called a superbasin.  In
Figure~\ref{fig:keyspace-zoomed}, superbasins are never re-visited because the
simulation only adds molecules; we can never reach a state with less molecules.
This is why keys are never re-accessed.  Despite these patterns, the following
characteristics of superbasins make it hard to detect them:

Detecting these superbasins can lead to more effective cache management
strategies because the height of the groups of key accesses is ``how much" of
the cache to evict and the width of the groups of key accesses is ``when" to
evict values from the cache.  The zoomed portion of
Figure~\ref{fig:keyspace-zoomed} shows how a single superbasin affects the key
accesses. Moving along the \(x\) axis shows that the number of unique keys
accessed over time grows while moving along the \(y\) axis shows that early
keys are accessed more often.

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that the following
techniques proliferate more parameters that need to be tuned per
hardware/software configuration. Furthermore, we find that many of the metrics
do not signal a new set of key accesses.  Below, we indicate with quotes which
parameters we need to add for each technique with and the value we find to work
best, via tuning and parameter sweeps, for one set of initial conditions.

@@

  \item Statistics: decay on each key counts down until 0; 0-valued keys are
evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed by large negative slope as signal for new set of key accesses.
``Zero-crossing", set to 40 seconds, for distance between small/large spikes to
avoid false positives; `window size", set to 200 seconds, for the size of the
moving average

@@

After trying these techniques we found that the basic algorithm in
Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
access patterns, which we call ``fans", by iterating backwards through the key
access trace, finding the lowest key ID, and comparing against the lowest key
ID we have seen so far (Line 7). We also maintain the top and bottom of each
group of key accesses (Line 13) so we can tell the ``how much" policy the
number of keys to evict (Line 19).  The algorithm iterates backwards over the
key access trace because a change in the minimum value signals a new group of
key accesses. No signal exists iterating left to right, as the maximum value
always increases and the minimum values at the bottom of each group of key
accesses are sparse.  For example, the maximum distance between values along
the bottom edge of the zoomed group of key accesses in
Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
between minimum values for the group of key accesses before is 0 seconds. As a
result of this sparseness, iterating left to right requires a ``window size"
parameter to determine when we think a minimum value will not show up again.
We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
number events, for simplicity because this approach avoids adding new
thresholds for key access pattern detection ({\it e.g.}, space between key
accesses, space between key IDs, and window size of consecutive key accesses).

### 2017-11-26 paper/sections/6_conclusion.tex
paper: addresses #45
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

State-of-the-art distributed file systems partition write-heavy workloads and
replicate read-heavy workloads, similar to the approach we are advocating
here.  IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients
write to different partitions by grabbing leases and caching ancestor metadata
for path traversal. ShardFS takes the replication approach to the extreme by
copying all directory state to all nodes. The Ceph file system
(CephFS)~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
techniques to a lesser extent; directories can be replicated or sharded but the
caching and replication policies are controlled with tunable parameters.  These
systems still need to be tuned by hand with {\it ad-hoc} policies designed for
specific applications.  Setting policies for migrations is arguably more
difficult than adding the migration mechanisms themselves.  For example,
IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
partitioning directories at a \emph{predefined} threshold. Mantle makes headway
in this space by providing a framework for exploring these policies, but does
not attempt anything more sophisticated (e.g., machine learning) to create
these policies.

@@

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Ultimately, we hope that this automation enables control of policies by
machines instead of administrators.

>>>>

State-of-the-art distributed file systems partition write-heavy workloads and
replicate read-heavy workloads, similar to the approach we are advocating
here.  IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients
write to different partitions by grabbing leases and caching ancestor metadata
for path traversal. ShardFS takes the replication approach to the extreme by
copying all directory state to all nodes. The Ceph file system
(CephFS)~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
techniques to a lesser extent; directories can be replicated or sharded but the
caching and replication policies are controlled with tunable parameters.  These
systems still need to be tuned by hand with {\it ad-hoc} policies designed for
specific applications.  Setting policies for migrations is arguably more
difficult than adding the migration mechanisms themselves.  For example,
IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
partitioning directories at a \emph{predefined} threshold. Mantle makes headway
in this space by providing a framework for exploring these policies, but does
not attempt anything more sophisticated ({\it e.g.}, machine learning) to create
these policies.

@@

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({\it e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Ultimately, we hope that this automation enables control of policies by
machines instead of administrators.
